---
title: "Avocados are too Popular"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("tidytext")
avo <- read_csv("/Users/gai7612/OneDrive/Documents/Data_Science_Project/avocados/avocado_app/avocados_conventional.csv")
```

## Background
</br><p>
Anyone else think that there is a huge obsession with avocados? Almost every brunch menu has avocado toast. So many restaurants offer fresh-made guacamole at your table. </br>
  And way too many people have been asking me if I want to add avocado to my burrito/sandwich/salad/pizza (pro tip: it costs extra!). </br>
  There are even <a href="https://www.ispot.tv/ad/wgAy/avocados-from-mexico-referee" target="_blank">commericials</a> for avocados. </br>
  I have been wondering why this has been happening. I have a few guesses: government subsidies, improvements in agriculture, a more health concious population, economic benefits for central/south America. </br>
  So, when I came across a dataset on <a href="https://tinyletter.com/data-is-plural/letters/data-is-plural-2018-09-26-edition", target="_blank">Avocado sales</a>, I had to take a look. </br></p>
</br>
<p>
Here is what the data look like</br>
</p>

```{r}
avo
```
</br>
And here is a simple interactive <a href="https://shiny.rstudio.com/", target="_blank">Shiny R</a> App for exploring linear relationships among price, volume, and types of avocados over time. 
<p>
  <iframe src = "https://golnaz.shinyapps.io/avocado_app/" , height = 500, width = 900>
  </iframe>
</p>
