---
title: "Avocados are too Popular"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("tidytext")
avo <- read_csv("/Users/golnazirgens/Box/Data Science Hobby/website/avocados/avocado_app/avocados_conventional.csv")
```

</br>

## Background
<p>
Avocados are everywhere lately. Almost every brunch menu has avocado toast. Many restaurants offer fresh-made guacamole at your table. And way too many people have been asking me if I want to add avocado to my burrito/sandwich/salad/pizza (pro tip: it costs extra!). There are even <a href="https://www.ispot.tv/ad/wgAy/avocados-from-mexico-referee" target="_blank">commericials</a> for avocados.</p> <p>
I have been wondering why this has been happening. I have a few guesses: government subsidies, improvements in agriculture, a more health concious population, economic benefits for central/south America. And I have also been wondering if it's just me or if avocados have truly gained popularity and if so, what did that gain look like? </br></p>
<p>
So, when I came across a dataset on <a href="https://tinyletter.com/data-is-plural/letters/data-is-plural-2018-09-26-edition", target="_blank">Avocado sales</a>, I had to take a look. 
</p>
</br>

## Initial Look
<p>
Here is what the data look like
</p>

```{r}
avo
```
</br>

## The Interactive App
<p>
And here is a simple interactive <a href="https://shiny.rstudio.com/", target="_blank">Shiny R</a> App for exploring linear relationships among price, volume, and types of avocados over time. 
</p>
<p>
  <iframe src = "https://golnaz.shinyapps.io/avocado_app/" , height = 500, width = 900>
  </iframe>
</p>
